---
title: "Download Prebuilt Demos"
vignette: >
  %\VignetteIndexEntry{Download Prebuilt Demos}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| include: false
library(shinyelectron)
```

Every demo is built with every runtime strategy on every supported platform and published as a release asset. Pick a row, check what it needs, and download.

What each strategy needs on the user's machine:

- `shinylive`: nothing. The app runs in WebAssembly.
- `bundled`: nothing. A runtime is embedded in the installer.
- `system`: R or Python already installed and on the PATH.
- `auto-download`: internet access on first launch to fetch the runtime.
- `container`: Docker or Podman installed.

```{r}
#| echo: false
#| results: asis
m <- shinyelectron:::demo_release_matrix()
base <- "https://github.com/coatless-rpkg/shinyelectron/releases/latest/download/"
cat("| Demo | Strategy | Platform | Requirement | Download |\n")
cat("|------|----------|----------|-------------|----------|\n")
for (i in seq_len(nrow(m))) {
  cat(sprintf("| %s | `%s` | %s %s | %s | [%s](%s%s) |\n",
    m$name[i], m$strategy[i], m$platform[i], m$arch[i],
    m$requirement[i], m$asset_name[i], base, m$asset_name[i]))
}
```

All links point at the latest release, so they stay current as new versions are published.
